### 2016-03-10 Projects/DelCo_2014_2015/Reports/Hose_Stream_Report/TeX_Files/DelCo_Hose_Stream_Tests.tex
Minor edits to hose stream report

\textit{Should I present flow rates in GPM only, or LPM (GPM)? What about the tables? As of now, GPM is used, but that goes against the rule of keeping the type of units consistent in a table.}
\begin{figure}[!ht]
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/monitor}
	\end{center} 
	\endminipage \hfill
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/handline}
	\end{center}
	\endminipage \hfill
	\caption[Monitor and handline equipped with a combination nozzle.]{Monitor (left) and 1.75~in handline (right) equipped with a combination nozzle and used to flow water during experiments.}
	\label{fig:monitor+handline}
\end{figure}
\FloatBarrier

@@

\subsection{Monitor Experiments}
\label{sec:monitor_procedure}
Three different tests were conducted using a monitor equipped with a combination nozzle to flow water at 120~GPM in a straight stream, narrow fog stream, and wide fog stream pattern. One test, Test 33, occurred in the East Structure, and the two other tests, Tests 16 and 17, occurred in the West Structure. A fourth test, Test 70, used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water at 180~GPM in a straight stream and solid stream, respectively. Because monitor nozzles are designed to flow water from a set position, water was always applied in a fixed position during the monitor experiments. Thus, the monitor experiments primarily focused on the impact of different hose stream patterns on air movement in a structure.

Test 33 utilized a monitor equipped with a combination nozzle to flow water from Room C of the East Structure to the area above the doorway on the south wall of Room B (Fig.~\ref{fig:test_33_pic}). A plan view of the structure's layout annotated with the approximate location of the bi-directional probes array used to measure gas velocity through an interior doorway (A6) and the general direction of the fully established flow path for Test 33 can be found in Fig.~\ref{fig:east_setup}. The procedure for Test 33 was the simplest of all the test procedures. First, water flowed in a straight stream pattern for 30 seconds, then the stream was changed to a narrow fog for 30 seconds, and finally to a wide fog stream for 30 seconds. The procedure was repeated an additional time.

@@

\begin{figure}[!ht]
	\minipage{2.15in}
	\begin{center}
		\includegraphics[width=2in]{../Figures/Pictures/SS_70}
	\end{center} 
	\endminipage
	\minipage{2.15in}
	\begin{center}
		\includegraphics[width=2in]{../Figures/Pictures/SB_70}
	\end{center}
	\endminipage
	\caption[Straight stream from combination nozzle and solid stream from smooth bore nozzle during Test 70.]{Straight stream (left) from combination nozzle and solid stream (right) from smooth bore nozzle during Test 70.}
	\label{fig:test_70_pic}
\end{figure}
\FloatBarrier

\subsection{Handline Experiments}
\label{sec:handline_procedure}
Numerous experiments were conducted during three different test series using a 1.75~in handline with a combination nozzle to flow water in a straight stream, narrow fog stream, and wide fog stream pattern. One test series, Test 34, occurred in the East Structure, and the remaining series, Tests 18 and 19, occurred in the West Structure. All three test series primarily focused on the impact of different hose stream and water application pattern selections on air movement in a structure.

Test 34 utilized a 1.75~in handline equipped with a combination nozzle to flow water from Room C of the East Structure into Room B (Fig.~\ref{fig:test_34_pic}). The layout of the structure was identical to the layout described above in Section~\ref{sec:monitor_procedure} and shown in Fig.~\ref{fig:east_setup}. The procedure for Test 34 began by flowing water in a straight stream aimed at the ceiling of Room B for 30 seconds in a fixed position. After 30 seconds, the nozzle was rotated in the clockwise direction for 30 seconds. Then, the nozzle was aimed at the south doorway of Room B and water was applied for 30 seconds in a fixed position followed by 30 seconds in a clockwise pattern. This process was repeated for narrow fog and wide fog streams. Then, the entire procedure was repeated for the three types of streams being applied in the fixed and counter-clockwise pattern.

@@

% Where was the handline aimed? Add flow chart? %
Tests 18 and 19 followed identical procedures that involved using a 1.75~in handline equipped with a combination nozzle to flow water from the north side double doors of the West Structure into the first floor room. For every experiment during the two tests, the water stream was aimed at the south side wall of the first floor. The difference between Tests 18 and 19 is that Test 18 contained the same flow path configuration presented in Fig.~\ref{fig:flow_path_1}, while Test 19 contained the same flow path configuration presented in Fig.~\ref{fig:flow_path_2}. Each test series began by using a straight stream to flow water in a fixed position. After 60 seconds of water flow, the stairwell door was opened. One minute later, the north side, west double door on the second floor was opened, and water continued to flow for 60 seconds. Next, the water flow was stopped, the two doors were closed, and the process was repeated three additional times for the sweeping, clockwise, and counterclockwise application patterns. This entire procedure was repeated using narrow fog and wide fog streams. Fig.~\ref{fig:test_18_pic} contains an image of a straight stream pattern being applied during Test 18 with the flow path fully established.

@@

\chapter{Results and Discussion}
\label{chap:results}
In the following sections, the measurements are presented in graphic and tabular form. In the graphs, an error bar will represent the estimated uncertainty of the measurement. In the tables, the uncertainty will be included in the caption of the table.

@@

Fig.~\ref{fig:Test_16_BDP_A10_Avg_All} contains a plot of the average measured velocity through the interior stairwell door (A10) for the straight, narrow fog, and wide fog stream patterns during Test 16. Similar plots of the average measured velocity through the interior doorway for Tests 17, 33, and 70 are presented in Appendix~\ref{chap:monitor_plots}.

@@

Excluding the straight stream during Test 33, there was significant air movement (gas velocity greater than 0.5 m/s) through the interior doorway while the flow path was fully established for every hose stream and target location combination during the monitor experiments. Table~\ref{table:all_mon_vel_avgs} lists the average velocity measured by the bi-directional probes array at the interior doorway when the flow path was fully established for each hose stream and target location combination studied during Tests 16, 17, 33, and 70.

@@

Looking at Tests 16, 17, and 33 in Table~\ref{table:all_mon_vel_avgs}, the narrow and wide fog streams caused higher velocities when the flow path was fully established through the interior doorway compared to the straight stream. This trend is also observed in Fig.~\ref{fig:Test_16_BDP_A10_Avg_All} above and Figs.~\ref{fig:Test_33_BDP_A6_Avg_All}~and~\ref{fig:Test_17_BDP_A10_Avg_All} in Appendix~\ref{chap:monitor_plots}. Additionally, the wide fog stream generally produced equal or higher air velocities compared to the narrow fog stream. Finally, the Test 70 results listed in Table~\ref{table:all_mon_vel_avgs} and plotted in Fig.~\ref{fig:Test_70_BDP_A10_Avg_All} suggest that a straight stream from a combination nozzle and a solid stream from a smooth bore nozzle produce approximately the same amount of air movement through the interior doorway.
\FloatBarrier

@@

Fig.~\ref{fig:Test_18_BDP_A10_Avg_All} contains a plot of the average measured velocity through the interior stairwell door (A10) for the straight, narrow fog, and wide fog stream patterns being applied in a fixed position, a sweeping pattern, a clockwise rotation, and counter clockwise rotation during Test 18. Similar plots of the average measured velocity through the interior doorway for Tests 19 and 34 are presented in Appendix~\ref{chap:handline_plots}.

@@

Tables~\ref{table:east_hand_A6_avgs}~and~\ref{table:west_hand_A10_avgs} list the average velocity through the interior doorway when the flow path was fully established for each hose stream and application pattern combination studied during Test 34 and Tests 18-19, respectively.

@@

Nearly every hose stream and application pattern combination using a narrow or wide fog stream during the handline experiments produced significant air movement (gas velocity greater than 0.5 m/s) through the interior doorway while the flow path was fully established. For the straight stream, however, significant air movement through the interior doorway was only reached once the stream was applied in a pattern that involved moving the handline. Similar to the monitor experiments, the average air velocity through the interior doorway while the flow path was established was always greater in magnitude for the application patterns using the narrow and wide fog streams than the same patterns using the straight stream.

Another important result that can be seen graphically in Figs.~\ref{fig:Test_18_BDP_A10_Avg_All} and \ref{fig:Test_19_BDP_A10_Avg_All} and in Tables~\ref{table:east_hand_A6_avgs} and \ref{table:west_hand_A10_avgs} is that there are no statistically significant differences in the amount of air flow through the doorway when rotating the hoseline in the clockwise direction compared to rotating it in the counter clockwise direction for any of three hose streams tested.

@@

Three experiments studied the differences between a straight stream, narrow fog stream, and wide fog stream emitted from a combination nozzle attached to a monitor aimed at specific locations in a structure for different ventilation configurations. For each of the three experiments, the average air velocity through the interior doorway while the flow path was fully established was greatest for the wide fog stream no matter the location and ventilation pattern within the structure. The magnitude of the average air velocity through the interior doorway while the flow path was fully established for the narrow fog stream was always equal to or greater than the magnitude of the average air velocity for the straight stream. Additionally, a fourth experiment using a monitor was conducted in which a straight stream from a combination nozzle was compared to a solid stream from a smooth bore nozzle aimed at a specified location for 60 seconds with a fully established flow path. The streams were compared for six different 60 seconds periods; three times for the monitor aimed at two locations. The results from the four monitor experiments suggest a wide fog stream impacts air movement within a structure the most out of any of the tested streams, followed by the narrow fog stream, and then the straight stream and solid stream from a smooth bore nozzle, which both caused approximately the same amount of air movement through the structure.

In addition to the monitor experiments, three different experiments were conducted using a handline to study the impact of different stream and application pattern combinations on air movement within a structure. Similar to the monitor experiments, the average air velocity through the interior doorway of a structure with a fully established flow path was greater for the wide fog and narrow fog streams than the straight stream. In fact, it was discovered that a straight stream only caused significant air movement through the interior doorway when it was applied in a moving pattern. Additionally, throughout all handline experiments, there were no statistically significant differences in the average air velocity measured through the interior doorway between the clockwise and counter clockwise application patterns.

>>>>

\textit{Should I present flow rates in GPM only, or LPM (GPM)? What about the tables? As of now, GPM is used, but that goes against the rule of keeping the type of units consistent in a table. Same goes for ``1.75 in'' handline.}
\begin{figure}[!ht]
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/monitor}
	\end{center} 
	\endminipage \hfill
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/handline}
	\end{center}
	\endminipage \hfill
	\caption[Monitor and handline equipped with a combination nozzle.]{Monitor (left) and 1.75~in handline (right) equipped with a combination nozzle and used to flow water during experiments.}
	\label{fig:monitor+handline}
\end{figure}
\FloatBarrier

@@

\subsection{Monitor Experiments}
\label{sec:monitor_procedure}
Three different tests were conducted using a monitor equipped with a combination nozzle to flow water at 120~GPM in a straight stream, narrow fog stream, and wide fog stream pattern. One test, Test 33, occurred in the East Structure, and the two other tests, Tests 16 and 17, occurred in the West Structure. A fourth test, Test 70, was conducted in the West Structure and used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water at 180~GPM in a straight stream and solid stream, respectively. Because monitors are designed to flow water from a set position, water was always applied in a fixed position during the monitor experiments. Thus, the monitor experiments primarily focused on the impact of different hose stream patterns on air movement in a structure.

Test 33 utilized a monitor equipped with a combination nozzle to flow water from Room C of the East Structure to the area above the doorway on the south wall of Room B (Fig.~\ref{fig:test_33_pic}). A plan view of the structure's layout annotated with the approximate location of the bi-directional probe array used to measure gas velocity through an interior doorway (A6) and the general direction of the fully established flow path for Test 33 can be found in Fig.~\ref{fig:east_setup}. The procedure for Test 33 was the simplest of all the test procedures. First, water was flowed in a straight stream pattern for 30 seconds, then the stream was changed to a narrow fog for 30 seconds, and finally to a wide fog stream for 30 seconds. The procedure was repeated an additional time.

@@

\begin{figure}[!ht]
	\minipage{2.15in}
	\begin{center}
		\includegraphics[width=2in]{../Figures/Pictures/SS_70}
	\end{center} 
	\endminipage
	\minipage{2.15in}
	\begin{center}
		\includegraphics[width=2in]{../Figures/Pictures/SB_70}
	\end{center}
	\endminipage
	\caption[Straight stream from combination nozzle and solid stream from smooth bore nozzle during Test 70.]{Straight stream flowing from the combination nozzle (left) and solid stream flowing from the smooth bore nozzle (right) during Test 70.}
	\label{fig:test_70_pic}
\end{figure}
\FloatBarrier

\subsection{Handline Experiments}
\label{sec:handline_procedure}
Three experiments were conducted using a 1.75~in handline with a combination nozzle to flow water in a straight stream, narrow fog stream, and wide fog stream pattern. One test, Test 34, occurred in the East Structure, and the remaining tests, Tests 18 and 19, occurred in the West Structure. All three tests primarily focused on the impact of different hose stream and water application pattern selections on air movement in a structure.

Test 34 utilized a 1.75~in handline equipped with a combination nozzle to flow water from Room C of the East Structure into Room B (Fig.~\ref{fig:test_34_pic}). The layout of the structure was identical to the layout for Test 33, described above in Section~\ref{sec:monitor_procedure} and presented in Fig.~\ref{fig:east_setup}. The procedure for Test 34 began by flowing water in a straight stream pattern aimed at the ceiling of Room B for 30 seconds in a fixed position. After 30 seconds, the nozzle was rotated in the clockwise direction for 30 seconds. Then, the nozzle was aimed at the south doorway of Room B, and water was applied for 30 seconds in a fixed position followed by 30 seconds in a clockwise pattern. This process was repeated for narrow fog and wide fog streams. Then, the entire procedure was repeated for the three types of streams being applied in the fixed and counterclockwise patterns.

@@

% Where was the handline aimed? Add flow chart? %
Tests 18 and 19 followed identical procedures that involved using a 1.75~in handline equipped with a combination nozzle to flow water from the north side double doors of the West Structure into the first floor room. The water stream was aimed at the south side wall of the first floor during the experiments. The difference between Tests 18 and 19 is that Test 18 contained the flow path configuration presented in Fig.~\ref{fig:flow_path_1}, while Test 19 contained the flow path configuration presented in Fig.~\ref{fig:flow_path_2}. Each test began by using a straight stream to flow water in a fixed position. After 60 seconds of water flow, the stairwell door was opened. One minute later, the north side, west double door on the second floor was opened, and water continued to flow for 60 seconds. Next, the water flow was stopped, the two doors were closed, and the process was repeated three additional times for the sweeping, clockwise, and counterclockwise application patterns. This entire procedure was repeated using narrow fog and wide fog streams. Fig.~\ref{fig:test_18_pic} contains an image of a straight stream pattern being applied during Test 18 with the flow path fully established.

@@

\chapter{Results and Discussion}
\label{chap:results}
In the following sections, the measurements are presented in graphic and tabular form. In the graphs, an error bar will represent the estimated uncertainty of the measurement. In the tables, the uncertainty will be included in the caption of the table as a percentage enclosed in brackets.

@@

Fig.~\ref{fig:Test_16_BDP_A10_Avg_All} contains a plot of the average measured velocity through the interior stairwell door (A10) for the straight, narrow fog, and wide fog stream patterns during Test 16. Similar plots of the average measured velocity through the interior doorway for Tests 17, 33, and 70 are presented in Appendix~\ref{chap:monitor_plots}. Additionally, Table~\ref{table:all_mon_vel_avgs} lists the average velocity measured by the bi-directional probe array at the interior doorway when the flow path was fully established for each hose stream and target location combination studied during Tests 16, 17, 33, and 70.

@@

@@

Looking at the table and figures, a number of trends can be observed. For example, there was significant air movement (gas velocity greater than 0.5 m/s) through the interior doorway while the flow path was fully established for every hose stream and target location combination except the straight stream in Test 33. Also, looking at Table~\ref{table:all_mon_vel_avgs}, the narrow and wide fog streams caused higher velocities through the interior doorway when the flow path was fully established compared to the straight stream. This trend is also observed above in Fig.~\ref{fig:Test_16_BDP_A10_Avg_All} and in Figs.~\ref{fig:Test_33_BDP_A6_Avg_All}~and~\ref{fig:Test_17_BDP_A10_Avg_All} in Appendix~\ref{chap:monitor_plots}. Furthermore, the wide fog stream generally produced equal or higher air velocities compared to the narrow fog stream. Finally, the Test 70 results listed in Table~\ref{table:all_mon_vel_avgs} and plotted in Fig.~\ref{fig:Test_70_BDP_A10_Avg_All} suggest that a straight stream from a combination nozzle and a solid stream from a smooth bore nozzle produce approximately the same amount of air movement through the interior doorway.
\FloatBarrier

@@

Fig.~\ref{fig:Test_18_BDP_A10_Avg_All} contains a plot of the average measured velocity through the interior stairwell door (A10) for the straight, narrow fog, and wide fog stream patterns being applied in a fixed position, a sweeping pattern, a clockwise rotation, and counterclockwise rotation during Test 18. Similar plots of the average measured velocity through the interior doorway for Tests 19 and 34 are presented in Appendix~\ref{chap:handline_plots}. Additionally, Tables~\ref{table:east_hand_A6_avgs}~and~\ref{table:west_hand_A10_avgs} list the average velocity measured by the bi-directional probe array at the interior doorway when the flow path was fully established for each hose stream and application pattern combination studied during Test 34 and Tests 18-19, respectively.

@@

@@

Nearly every hose stream and application pattern combination using a narrow or wide fog stream during the handline experiments produced significant air movement (gas velocity greater than 0.5 m/s) through the interior doorway while the flow path was fully established. For the straight stream, however, significant air movement through the interior doorway was only reached once the stream was applied in a pattern that involved moving the handline. Similar to the monitor experiments, the average air velocity through the interior doorway when the flow path was established was always greater in magnitude for the application patterns using the narrow and wide fog streams than the same patterns using the straight stream. Another important result that can be seen graphically in Figs.~\ref{fig:Test_18_BDP_A10_Avg_All} and \ref{fig:Test_19_BDP_A10_Avg_All} and in Tables~\ref{table:east_hand_A6_avgs} and \ref{table:west_hand_A10_avgs} is that there are no statistically significant differences in the amount of air flow through the doorway when rotating the hoseline in the clockwise direction compared to rotating it in the counterclockwise direction for any of three hose streams tested.

@@

Three experiments studied the differences between a straight stream, narrow fog stream, and wide fog stream emitted from a combination nozzle attached to a monitor aimed at specific locations in a structure for different ventilation configurations. For each of the three experiments, the average air velocity through the interior doorway while the flow path was fully established was greatest for the wide fog stream no matter the location and ventilation pattern within the structure. The magnitude of the average air velocity through the interior doorway while the flow path was fully established for the narrow fog stream was always equal to or greater than the magnitude of the average air velocity for the straight stream. Additionally, a fourth experiment using a monitor was conducted in which a straight stream from a combination nozzle was compared to a solid stream from a smooth bore nozzle aimed at a specified location for 60 seconds with a fully established flow path. The streams were compared for six different 60 seconds periods during which the flow path was fully established: three times each for the monitor aimed at two locations. The results from the four monitor experiments suggest a wide fog stream impacts air movement within a structure the most out of any of the tested streams, followed by the narrow fog stream, and then the straight stream and solid stream from a smooth bore nozzle, which both caused approximately the same amount of air movement through the structure.

In addition to the monitor experiments, three different experiments were conducted using a handline to study the impact of different stream and application pattern combinations on air movement within a structure. Similar to the monitor experiments, the average air velocity through the interior doorway of a structure with a fully established flow path was greater for the wide fog and narrow fog streams than the straight stream. In fact, it was discovered that a straight stream only caused significant air movement through the interior doorway when it was applied in a moving pattern. Additionally, throughout all handline experiments, there were no statistically significant differences in the average air velocity measured through the interior doorway between the clockwise and counterclockwise application patterns.
